---
title: "Detecting Game Framing in the Danish Parliament"
author: "Bjørn August Skeel-Gjørling"
date: "14 apr 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import libraries
```{r, message=FALSE, warning=F}
#Loading packages
library(plyr)
library(dplyr)
library(readr)
library(ggplot2)
library(knitr)
library(scales)
library(zoo)
library(lubridate)
```

## Import data
In the code section, I import the dataset analysis in Python. Variables are renamed and duplicates are dropped. Furthermore, the information on the date of the debates are converted into a date format using lubridate
```{r, warning=FALSE}
game_frame_data = read.csv("https://raw.githubusercontent.com/basgpol/ps239t-final-project/master/data/debate_game_frame.csv", header = TRUE)

game_frame_data = read.csv("debate_game_frame.csv", header = TRUE) %>% 
  rename(game_frame_proportion = Game.frame.proportion) %>%   #renaming variable
  mutate(date = ymd_hms(game_frame_data$Time)) %>%  #converting to date format using lubridate
  select(c(4,3)) %>%  # dropping the first index variable
  distinct(date, game_frame_proportion) #dropping duplicates

```

## Preprocessing the data
In the following chunk of code, I first create a vecter with the lastest three parliamentary elections in Denmark. Thereafter, I create three new variables to the dataset. The first three calculates the distance in days to the election in 2015, 2011 and 2007 respectively. Transcripts after the election set to NA.
```{r, warning=FALSE}
# Time to Election dates
election_dates = c(ymd_hms("2015-06-18T09:00:00"), ymd_hms("2011-09-15T09:00:00"), ymd_hms("2007-11-13T09:00:00")) # creating a list of election dates

game_frame_data_dates = game_frame_data %>% 
  mutate(time_to_2015_election = election_dates[1] - game_frame_data$date) %>% # create a variable that calculates distance in days to 2015 election 
  mutate(time_to_2015_election = replace(time_to_2015_election, which(time_to_2015_election<0), NA)) %>% #replace negatives (after election) to NA
  mutate(time_to_2011_election = election_dates[2] - game_frame_data$date) %>% # create a variable that calculates distance in days to 2011 election 
  mutate(time_to_2011_election = replace(time_to_2011_election, which(time_to_2011_election<0), NA)) %>%  #replace negatives (after election) to NA 
  mutate(time_to_2007_election = election_dates[3] - game_frame_data$date) %>% # create a variable that calculates distance in days to 2011 election 
  mutate(time_to_2007_election = replace(time_to_2007_election, which(time_to_2007_election<0), NA)) #replace negatives (after election) to NA
```

I thereafter creates a loop that returns a vector with number of days to next election no matter what election we are talking about. This vector is included as a new variable in the dataset. 

```{r, warning=FALSE}
# For loop returning an vector with the days to next election
days_to_next_election_vetor = c() # create an empty vector
for (index in array(1:nrow(game_frame_data_dates))){
  days_to_next_election_vetor = c(days_to_next_election_vetor, min(c(game_frame_data_dates$time_to_2015_election[index], 
        game_frame_data_dates$time_to_2011_election[index],
        game_frame_data_dates$time_to_2007_election[index]), na.rm = TRUE))
}
# Replace inf (observation since latest election in 2015) with NA
days_to_next_election_vetor[days_to_next_election_vetor == Inf] <- NA

#Adding a extra variable (days to next election) to the dataset 
game_frame_final = game_frame_data_dates %>% 
  mutate(days_to_election = days_to_next_election_vetor)  
```

## Visualizations

### The cyclical nature of game framing

I first make a time-line of the usage of game frame words since end 2007. The large outliers make it difficult to detect a clear trend.

```{r, message=FALSE }
### Overall time trend
overall_trend = ggplot(game_frame_final, aes(x = date, y = game_frame_proportion*100)) +
  geom_point() + # create a plot
  geom_smooth() + # create a smooth line
  theme_minimal() + # set theme
  xlab("Date")+ # set label for x-axis
  ylab("Percentage of game frame words")+ # set label for y-axis
  theme(plot.title = element_text(hjust = 0.5)) + # center plot title
  ggtitle("Overall time trend of game framing") # plot title

overall_trend
```

In the next cell, I remove three outliers of the dataset. That is the three transcripts, where the percentage of game frame words is above 2 percentage


```{r, message=FALSE, warning=FALSE}
game_frame_no_outliers = game_frame_final[!rowSums(game_frame_data[2] >0.02),]

overall_trend = ggplot(game_frame_no_outliers, aes(x = date, y = game_frame_proportion*100)) +
  geom_point() + # create a plot
  geom_smooth() + # create a smooth line
  theme_minimal() + # set theme
  xlab("Date")+ # set label for x-axis
  ylab("Percentage of game frame words")+ # set label for y-axis
  theme(plot.title = element_text(hjust = 0.5)) + # center plot title
  ggtitle("Overall time trend of game framing without outliers") # plot title

overall_trend

```

There is create variation in the presence of game frame words. I create a varaiable that calculates the rolling average of the percentage of game frame words. The time trend becomes much more clear.

```{r, message=FALSE, warning=FALSE}
game_frame_data_rolling_mean = game_frame_no_outliers %>% 
  mutate(rolling_game_frame_average = rollmean(game_frame_proportion, k = 5, na.pad = TRUE))
  

overall_trend = ggplot(game_frame_data_rolling_mean, aes(x = date, y = rolling_game_frame_average*100)) +
  geom_point() + # create a plot
  geom_smooth() + # create a smooth line
  theme_minimal() + # set theme
  xlab("Date")+ # set label for x-axis
  ylab("Rolling mean of game frame words (percentage)")+ # set label for y-axis
  theme(plot.title = element_text(hjust = 0.5)) + # center plot title
  ggtitle("Overall time trend of game framing (rolling average)") # plot title

overall_trend
```

It is clear that the presence of game framing undergo a cyclical trend. I include lines to the plot representing elections dates.Just as expected, the presence of game framing tops simultaneously with the parliamentary elections. 

```{r, message=FALSE, warning=FALSE}
overall_trend = ggplot(game_frame_data_rolling_mean, aes(x = date, y = rolling_game_frame_average*100)) +
  geom_point() + # create a plot
  geom_smooth() + # create a smooth line
  geom_vline(aes(xintercept = as.numeric(election_dates[1])), colour="red") +
  geom_vline(aes(xintercept = as.numeric(election_dates[2])), colour="red") +
  geom_vline(aes(xintercept = as.numeric(election_dates[3])), colour="red") +
  theme_minimal() + # set theme
  xlab("Date")+ # name label for x-axis
  ylab("Rolling mean of game frame words (percentage)")+ # name label for y-axis
  theme(plot.title = element_text(hjust = 0.5)) + # center plot title
  ggtitle("Overall time trend of game framing (rolling average)") # name title

overall_trend
```

### Game framing an distance to next election
The next step is my analysis is to investigate if we can find an association between time to next parliamentary election and the usage of game framing in parliamentary debates. The first scatter plot with a linear regression line shows no sign of an association. 

```{r, message=FALSE, warning=FALSE}
# Association between days to election and presence of game frame

days_to_election_plot = ggplot(game_frame_data_rolling_mean, aes(x = days_to_election, y = game_frame_proportion*100)) +
  geom_point() + # create a plot
  geom_smooth(method='lm', formula = y~x) + # create a smooth line
  theme_minimal() + # set theme
  xlab("Days to next election")+ # name label for x-axis
  ylab("Percentage of game frame words")+ # name label for y-axis
  theme(plot.title = element_text(hjust = 0.5)) + # center plot title
  ggtitle("Time to election and game framing") # name title

days_to_election_plot
```

If we instead measure the presence of game framing as a rolling averages, we see a quite different picture. Suddenly, it seams like the presence of game framing increases as there is longer time to next election. This is opposite of what we would expect. 
```{r, message=FALSE, warning=FALSE}
# Association between days to election and presence of game frame as a rolling average

days_to_election_plot = ggplot(game_frame_data_rolling_mean, aes(x = days_to_election, y = rolling_game_frame_average*100)) +
  geom_point() + # create a plot
  geom_smooth(method='lm', formula = y~x) + # create a smooth line
  theme_minimal() + # set theme
  xlab("Days to next election")+ # name label for x-axis
  ylab("Rolling mean of game frame words (percentage)")+ # name label for y-axis
  theme(plot.title = element_text(hjust = 0.5)) + # center plot title
  ggtitle("Time to election and game framing as rolling average") # name title

days_to_election_plot
```

We saw the the illustration of the cyclical trend of the game framing that the presence of game frame words is high both before and after an election. Therefore, it makes sence to measure on *days to closests election* rather than *days to next elction*. In the following code section it create a dataset that contains that exact variable. 

```{r, message=FALSE, warning=FALSE}
### Creating the 

# Creating new dataset measuring days to closests election instead of days to next election
closests_election_data = game_frame_data %>% 
  mutate(time_to_2015_election = abs(election_dates[1] - game_frame_data$date)) %>% # create a variable that calculates distance in days to 2015 election 
  mutate(time_to_2011_election = abs(election_dates[2] - game_frame_data$date)) %>% # create a variable that calculates distance in days to 2011 election 
  mutate(time_to_2007_election = abs(election_dates[3] - game_frame_data$date)) # create a variable that calculates distance in days to 2011 election 

# For loop returning an vector with the days to closets election
days_to_closets_election_vetor = c() # create an empty vector
for (index in array(1:nrow(closests_election_data))){
  days_to_closets_election_vetor = c(days_to_closets_election_vetor, min(c(closests_election_data$time_to_2015_election[index], 
                                                                           closests_election_data$time_to_2011_election[index],
                                                                           closests_election_data$time_to_2007_election[index]), na.rm = TRUE))
}

#Adding a extra variable (days to next election) to the dataset 
closests_election_data = closests_election_data %>% 
  mutate(days_to_closets_election = days_to_closets_election_vetor) %>% #add the new variable
  mutate(rolling_game_frame_average = rollmean(game_frame_proportion, k = 5, na.pad = TRUE)) # create a variable measuring the rolling average of game frame words
  
```

If we now plot the days to closets election with the presence of game framing with a linear regression line, we find what we expected: the exact opposite of the previous plot. The closer a parliamentary debates is to the closets election, the more does the politicians use game frame words in the parliamentary debate. 

```{r, message=FALSE, warning=FALSE}
# Association between days to election and presence of game frame as a rolling average

days_to_election_closets_plot = ggplot(closests_election_data, aes(x = days_to_closets_election, y = rolling_game_frame_average*100)) +
  geom_point() + # create a plot
  geom_smooth(method='lm', formula = y~x) + # create a smooth line
  theme_minimal() + # set theme
  xlab("Days to closests election")+ # name label for x-axis
  ylab("Rolling mean of game frame words (percentage)")+ # name label for y-axis
  theme(plot.title = element_text(hjust = 0.5)) + # center plot title
  ggtitle("Time to closests election and game framing as rolling average") # name title

days_to_election_closets_plot
```

Furthermore, we find a statistically significant negative association between days to closests election and rolling average of game frame words in the debate. In other words, the more days there is to an election (both previous or next election) the less frequent do politicians use game frame words in the parliamentary debates. This is statistically significant with a t-value of -9.34.  

```{r}
ln_model = lm(rolling_game_frame_average ~ days_to_closets_election, data=closests_election_data)

summary(ln_model)
```




